---
title: "Meeting R"
output:
  learnr::tutorial:
      progressive: true
runtime: shiny_prerendered
description: >
  Learn the basics of R: interpreter, expressions, evaluation and variables.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Interpreter and expressions

We'll see here how we can write R expressions and get them executed.


### R as a calculator 

Write the R code required to compute two plus two:

```{r two-plus-two, exercise=TRUE}

```

```{r two-plus-two-solution}
2+2
```

### Great! You made it!
Do you see something strange? Well, what about the [1] in the output? Keep it in the back of your head!


### Hello world!

Write the R code to print "Hello world!".

```{r hello, exercise=TRUE}
print('miao')
```

```{r hello-solution}
print('Hello world!')
```

### Your first function
`print` is an R function, provided by the `base` package. More on functions and packages in the theory.


## Variables

### Using variables

Variables are a basic element of every programming language. Try to use two variables to add 2 with 3.

```{r variables, exercise=TRUE}
a = 2
b = 6
a * b
```

```{r variables-solution}
a = 2
b = 3
a + b
```

### Varying variables

This things are called variables for a reason, right? Try to make `a` evaluate to 6.

```{r change-variables, exercise=TRUE}
a = 2
print(a)

# do stuff

# a should now be 6
print(a)
```

```{r change-variables-solution}
a = 2
print(a)

a = 6
print(a)
```

## Types

Types are a fundamental part of every programming language. R adopts dynamic typing, meaning that types are inferred by the language and the programmer does not need to make them explicit.

### Base types

In the following snippet you find five variables of different types. Write the code to print the type of each.

```{r types, exercise=TRUE}
a = 2
b = 'miao'
c = TRUE
d = 3.14
e = 2+2i

```

```{r types-solution}
a = 2
b = 'miao'
c = TRUE
d = 3.14
e = 2+2i

print(class(a))
print(class(b))
print(class(c))
print(class(d))
print(class(e))
```

### Numeric types 1
The variables `a` and `d` share the same type. Actually, the same high-level type. Do they share also the same low-level type? Check it below.
```{r numeric-types-1, exercise=TRUE}
a = 2
d = 3.14

```

```{r numeric-types-1-solution}
a = 2
d = 3.14

print(typeof(a))
print(typeof(d))
```

### Numeric types 2

Yes, looks like they do. Change the code to make `a` an integer.

```{r numeric-types-2, exercise=TRUE}
a = 2

print(typeof(a))
```

```{r numeric-types-2-solution}
a = 2L
print(typeof(a))
```

### Checking types 1

How would you check if the following variable is numeric?.

```{r checking-types-1, exercise=TRUE}
a = 2
```

```{r checking-types-1-solution}
a = 2
is.numeric(a)
```

### Checking types 2

How would you check if the following variable is an integer?.

```{r checking-types-2, exercise=TRUE}
a = 2
```

```{r checking-types-2-solution}
a = 2
is.integer(a)
```
