---
title: "Meeting R"
output:
  learnr::tutorial:
      progressive: true
runtime: shiny_prerendered
description: >
  Learn the basics of R: interpreter, expressions, evaluation and variables.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Interpreter and expressions

We'll see here how we can write R expressions and get them executed.


### R as a calculator 

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

```{r two-plus-two-solution}
2+2
```

### Great! You made it!
Do you see something strange? Well, what about the [1] in the output? Keep it in the back of your head!


### Hello world!

Write the R code to print "Hello world!".

```{r hello, exercise=TRUE}
print('miao')
```

```{r hello-solution}
print('Hello world!')
```

### Your first function
`print` is an R function, provided by the `base` package. More on functions and packages in the theory.


## Variables

### Using variables

Variables are a basic element of every programming language. Try to use two variables to add 2 with 3.

```{r variables, exercise=TRUE}
a = 2
b = 6
a * b
```

```{r variables-solution}
a = 2
b = 3
a + b
```

### Varying variables

This things are called variables for a reason, right? Try to make `a` evaluate to 6.

```{r change-variables, exercise=TRUE}
a = 2
print(a)

# do stuff

# a should now be 6
print(a)
```

```{r change-variables-solution}
a = 2
print(a)

a = 6
print(a)
```
