---
title: "Data Structures"
output:
  learnr::tutorial:
      progressive: true
runtime: shiny_prerendered
description: >
  Use vectors, lists, and dataframes: the most common data structures in R.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Vectors

Vectors are the simplest data structure. In R, also scalar variables are treated as vectors.


### Creating vectors

Let us create some vector. For instance, create a vector `int_vec` of two integers and a vector `char_vec` of two strings.

```{r vect-1, exercise=TRUE}




```

```{r vect-1-solution}
int_vec = c(1L, 2L)
char_vec = c('Miao', 'Bau')
```

### Creating sequences

Many R functions return vectors. For instance, let us create a vector of 100 equally spaced real numbers between 1 and 10.

<div class="alert alert-info">
  **Hint** You already know it. Google is your friend. But so I am: try running `?seq`.
</div>


```{r vect-2, exercise=TRUE}



```

```{r vect-2-solution}
x = seq(1, 10, length.out = 100)
x
```

### Vector operations 1

R supports natively operations between numeric vectors. For example, if you add two vectors, elements get summed up element-wise. Compute and print the element-wise sum and multiplication of the following vectors.

```{r vect-3, exercise=TRUE}
x = c(1, 2, 3)
y = c(9, 8, 7)

```

```{r vect-3-solution}
x = c(1, 2, 3)
y = c(9, 8, 7)
print(x+y)
print(x*y)
```

### Vector operations 2

Same as before. But, wait: before running the code, what do you expect?

```{r vect-4, exercise=TRUE}
x = c(1, 2, 3, 4, 5)
y = c(9, 8)

```

```{r vect-4-solution}
x = c(1, 2, 3, 4, 5)
y = c(9, 8)
print(x+y)
print(x*y)
```

### Functions on vectors

Several R functions work on vectors. Compute and print the mean, standard deviation, max and min of the following vector.

```{r vect-5, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r vect-5-solution}
x = c(1, 5, 7, 2, 4)
print(mean(x))
print(sd(x))
print(max(x))
print(min(x))
```

## Indexing vectors

Indexing is the operation we perform to access specific elements in the vector. There are several ways of indexing a vector in R, we will experiment some of them.

### Indexing 1

Using positive integers as an index, extract and print a vector containing 5 and 2 from the following vector.

```{r index-1, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-1-solution}
x = c(1, 5, 7, 2, 4)
print(x[c(2, 4)])
```

### Indexing 2

Using negative integers as an index, extract and print a vector containing 5 and 2 from the following vector.

```{r index-2, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-2-solution}
x = c(1, 5, 7, 2, 4)
print(x[- c(1, 3, 5)])
```

### Indexing 3

Using logical values as an index, extract and print a vector containing 5 and 2 from the following vector.

```{r index-3, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-3-solution}
x = c(1, 5, 7, 2, 4)
print(x[c(F, T, F, T, F)])
```

### Indexing 4

Extract and print the first three elements of the following vector.

```{r index-4, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-4-solution}
x = c(1, 5, 7, 2, 4)
print(x[1:3])
```

### Indexing 5

Extract and print the last three elements of the following vector.

```{r index-5, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-5-solution}
x = c(1, 5, 7, 2, 4)
print(x[3:length(x)])
```

## Lists

Vectors are made of elements of the same type, lists are not. We'll play with lists in the next exercises.

### Creating lists

Let us create a list, where element `a` equals `'miao'`, element `b` equals `2` and element `c` equals a vector made of 1 and 3.

```{r list-1, exercise=TRUE}
x = list() # complete this line
x
```

```{r list-1-solution}
x = list(a = 'miao', b = 2, c = c(1, 3)) # complete this line
x
```

### Indexing lists 1

Extract from the following list the element `b`.

```{r list-2, exercise=TRUE}
x = list(a = 'miao', b = 2, c = c(1, 3))

```

```{r list-2-solution}
x = list(a = 'miao', b = 2, c = c(1, 3))
x$b
```

### Indexing lists 2

Extract from the following list the element `b`, returning a list.

```{r list-3, exercise=TRUE}
x = list(a = 'miao', b = 2, c = c(1, 3))

```

```{r list-3-solution}
x = list(a = 'miao', b = 2, c = c(1, 3))
x['b']
```

### Indexing lists 3

Extract from the following list the element `3` of the element `c`.

```{r list-4, exercise=TRUE}
x = list(a = 'miao', b = 2, c = c(1, 3))

```

```{r list-4-solution}
x = list(a = 'miao', b = 2, c = c(1, 3))
c_vect = x$c
c_vect[2]
```

## Data frames

Data frames are the primary structure to store data sets. We'll see here only basic operations, but we'll use extensively data frames and their evolution, tibbles, in the next meetings.

We'll use the `mpg` dataset. These datasets appear frequently in R examples. 

The readings in this tutorial follow [_R for Data Science_](http://r4ds.had.co.nz/), sections 3.2 and 5.1. 

### What is a data frame?

A __data frame__ is a rectangular collection of values, usually organized so that variables appear in the columns and observations appear in rows. 

Here is an example: the `mpg` data frame contains observations collected by the US Environmental Protection Agency on 38 models of cars. To see the `mpg` data frame, type `mpg` in the code chunk below.


```{r mpg, exercise = TRUE}
library(ggplot2)
mpg = as.data.frame(mpg)

```

```{r mpg-solution}
library(ggplot2)
mpg = as.data.frame(mpg)
mpg
```

### Dataset documentation

Did you notice how much information was inside `mpg`? Me too. Sometimes the contents of a data frame do not fit on a single screen, which makes them difficult to inspect. We'll look at an alternative to using and examining data frames soon. But first let's get some help...

You can learn more about `mpg` by opening its help page. The help page will explain where the `mpg`dataset comes from and what each variable in `mpg` describes. To open the help page, type `?mpg` in the code chunk below and then click "Submit Answer".

```{r help, exercise = TRUE}


```



```{r help-solution}
?mpg
```

### ? syntax

You can open a help page for any object that comes with R or with an R package. To open the help page, type a `?` before the object's name and then run the command, as you did with `?mpg`. This technique works for functions, packages, and more. 

Notice that objects created by you or will not have a help page (unless you make one).

### Exercises

Use the code chunk below to answer the following questions.

```{r exercises1, exercise = TRUE}
```

```{r quiz1, echo = FALSE}
quiz(caption = "Quiz",
  question("In `mpg`, what it the type of the column `mpg`?",
           answer("character"), 
           answer("numeric", correct = TRUE),
           answer("integer", correct = TRUE),
           answer("double"),
           incorrect = "Incorrect.\nHint: you can extract the column from the dataframe and use `is.character`, `is.numeric` and similar to test the type.", 
           allow_retry = TRUE
           
  ),
  question("In `mpg`, what it the average engine displacement?",
           answer("2.390618"),
           answer("3.471795", correct = TRUE), 
           answer("4.421309"),
           answer("It cannot be computed"),
           incorrect = "Incorrect.\nHint: the columns of a dataframe are vectors, so you can compute the mean.", 
           allow_retry = TRUE
           
  ),
  question("In `mpg`, what is the most common number of cylinders?",
           answer("4", correct = TRUE), 
           answer("5"),
           answer("6"),
           answer("8"),
           incorrect = "Incorrect.\nHint: what about the `table` function on the `cyl` column?", 
           allow_retry = TRUE
           
  ),
  question("What does the `drv` variable of `mpg` describe?",
           answer("Whether or not the vehicle has driver side airbags"), 
           answer("Whether a car is automatic or manual transmission"),
           answer("The number of cylinders in the car's engine"),
           answer("Something else", correct = TRUE, message = "`drv` describes the type of drivetrain in a car: front wheel drive, rear wheel drive, or four wheel drive."), 
           allow_retry = TRUE
  ),
  question("How many rows are in the data frame named `cars`?",
           answer("2"), 
           answer("25"),
           answer("50", correct = TRUE),
           answer("100"),
           incorrect = "Incorrect.\nHint: R numbers the rows of a data frame when it displays the contents of a data frame. As a result, you can spot the number of rows in `cars` by examining `cars` in the code block above.", 
           allow_retry = TRUE
  ),
  question("How many columns are in the data frame named `cars`?",
           answer("1"), 
           answer("2", correct = TRUE),
           answer("4"),
           answer("more than four"),
           incorrect = "Incorrect.\nHint: If you inspect the contents of `cars` in the code block above, it should be pretty easy to count the number of columns.", 
           allow_retry = TRUE
  )
)
```
