---
title: "Data Structures"
output:
  learnr::tutorial:
      progressive: true
runtime: shiny_prerendered
description: >
  Use vectors, lists, and dataframes: the most common data structures in R.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Vectors

Vectors are the simplest data structure. In R, also scalar variables are treated as vectors.


### Creating vetors

Let us create some vector. For instance, create a vector `int_vec` of two integers and a vector `char_vec` of two strings.

```{r vect-1, exercise=TRUE}




```

```{r vect-1-solution}
int_vec = c(1L, 2L)
char_vec = c('Miao', 'Bau')
```

### Creating sequences

Many R functions return vectors. For instance, let us create a vector of 100 equally spaced real numbers between 1 and 10.

<div class="alert alert-info">
  **Hint** You already know it. Google is your friend. But so I am: try running `?seq`.
</div>


```{r vect-2, exercise=TRUE}



```

```{r vect-2-solution}
x = seq(1, 10, length.out = 100)
x
```

### Vector operations 1

R supports natively operations between numeric vectors. For example, if you add two vectors, elements get summed up element-wise. Compute and print the element-wise sum and multiplication of the following vectors.

```{r vect-3, exercise=TRUE}
x = c(1, 2, 3)
y = c(9, 8, 7)

```

```{r vect-3-solution}
x = c(1, 2, 3)
y = c(9, 8, 7)
print(x+y)
print(x*y)
```

### Vector operations 2

Same as before. But, wait: before running the code, what do you expect?

```{r vect-4, exercise=TRUE}
x = c(1, 2, 3, 4, 5)
y = c(9, 8)

```

```{r vect-4-solution}
x = c(1, 2, 3, 4, 5)
y = c(9, 8)
print(x+y)
print(x*y)
```

### Functions on vectors

Several R functions work on vectors. Compute and print the mean, standard deviation, max and min of the following vector.

```{r vect-5, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r vect-5-solution}
x = c(1, 5, 7, 2, 4)
print(mean(x))
print(sd(x))
print(max(x))
print(min(x))
```
