---
title: "Data Structures"
output:
  learnr::tutorial:
      progressive: true
runtime: shiny_prerendered
description: >
  Use vectors, lists, and dataframes: the most common data structures in R.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Vectors

Vectors are the simplest data structure. In R, also scalar variables are treated as vectors.


### Creating vectors

Let us create some vector. For instance, create a vector `int_vec` of two integers and a vector `char_vec` of two strings.

```{r vect-1, exercise=TRUE}




```

```{r vect-1-solution}
int_vec = c(1L, 2L)
char_vec = c('Miao', 'Bau')
```

### Creating sequences

Many R functions return vectors. For instance, let us create a vector of 100 equally spaced real numbers between 1 and 10.

<div class="alert alert-info">
  **Hint** You already know it. Google is your friend. But so I am: try running `?seq`.
</div>


```{r vect-2, exercise=TRUE}



```

```{r vect-2-solution}
x = seq(1, 10, length.out = 100)
x
```

### Vector operations 1

R supports natively operations between numeric vectors. For example, if you add two vectors, elements get summed up element-wise. Compute and print the element-wise sum and multiplication of the following vectors.

```{r vect-3, exercise=TRUE}
x = c(1, 2, 3)
y = c(9, 8, 7)

```

```{r vect-3-solution}
x = c(1, 2, 3)
y = c(9, 8, 7)
print(x+y)
print(x*y)
```

### Vector operations 2

Same as before. But, wait: before running the code, what do you expect?

```{r vect-4, exercise=TRUE}
x = c(1, 2, 3, 4, 5)
y = c(9, 8)

```

```{r vect-4-solution}
x = c(1, 2, 3, 4, 5)
y = c(9, 8)
print(x+y)
print(x*y)
```

### Functions on vectors

Several R functions work on vectors. Compute and print the mean, standard deviation, max and min of the following vector.

```{r vect-5, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r vect-5-solution}
x = c(1, 5, 7, 2, 4)
print(mean(x))
print(sd(x))
print(max(x))
print(min(x))
```

## Indexing vectors

Indexing is the operation we perform to access specific elements in the vector. There are several ways of indexing a vector in R, we will experiment some of them.

### Indexing 1

Using positive integers as an index, extract and print a vector containing 5 and 2 from the following vector.

```{r index-1, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-1-solution}
x = c(1, 5, 7, 2, 4)
print(x[c(2, 4)])
```

### Indexing 2

Using negative integers as an index, extract and print a vector containing 5 and 2 from the following vector.

```{r index-2, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-2-solution}
x = c(1, 5, 7, 2, 4)
print(x[- c(1, 3, 5)])
```

### Indexing 3

Using logical values as an index, extract and print a vector containing 5 and 2 from the following vector.

```{r index-3, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-3-solution}
x = c(1, 5, 7, 2, 4)
print(x[c(F, T, F, T, F)])
```

### Indexing 4

Extract and print the first three elements of the following vector.

```{r index-4, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-4-solution}
x = c(1, 5, 7, 2, 4)
print(x[1:3])
```

### Indexing 5

Extract and print the last three elements of the following vector.

```{r index-5, exercise=TRUE}
x = c(1, 5, 7, 2, 4)


```

```{r index-5-solution}
x = c(1, 5, 7, 2, 4)
print(x[3:length(x)])
```

## Lists

Vectors are made of elements of the same type, lists are not. We'll play with lists in the next exercises.

### Creating lists

Let us create a list, where element `a` equals `'miao'`, element `b` equals `2` and element `c` equals a vector made of 1 and 3.

```{r list-1, exercise=TRUE}
x = list() # complete this line
x
```

```{r list-1-solution}
x = list(a = 'miao', b = 2, c = c(1, 3)) # complete this line
x
```

### Indexing lists 1

Extract from the following list the element `b`.

```{r list-2, exercise=TRUE}
x = list(a = 'miao', b = 2, c = c(1, 3))

```

```{r list-2-solution}
x = list(a = 'miao', b = 2, c = c(1, 3))
x$b
```

### Indexing lists 2

Extract from the following list the element `b`, returning a list.

```{r list-3, exercise=TRUE}
x = list(a = 'miao', b = 2, c = c(1, 3))

```

```{r list-3-solution}
x = list(a = 'miao', b = 2, c = c(1, 3))
x['b']
```

### Indexing lists 3

Extract from the following list the element `3` of the element `c`.

```{r list-4, exercise=TRUE}
x = list(a = 'miao', b = 2, c = c(1, 3))

```

```{r list-4-solution}
x = list(a = 'miao', b = 2, c = c(1, 3))
c_vect = x$c
c_vect[2]
```
