---
title: "Meeting R"
output:
  learnr::tutorial:
      progressive: true
runtime: shiny_prerendered
description: >
  Learn the basics of R: interpreter, expressions, evaluation and variables.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Functions

Functions are reusable chunks of code which perform a specific operation. Many useful functions are defined in existing R packages, but you can always write your own.


### Base functions

Many useful functions have already been implemented. No need to reinvent the wheel! Try printing the variables in the following chunk using the `print` function of the `base` package.

```{r print, exercise=TRUE}
a = 1
b = c(1.1, 2.2, 3.3)
l = list(a='miao', b=2)

```

```{r print-solution}
a = 1
b = c(1.1, 2.2, 3.3)
l = list(a='miao', b=2)

print(a)
print(b)
print(l)
```

### Plotting

Also, R has great functions for plotting. Use the `plot` function in the `base` package to perform a line plot of x versus y.

<div class="alert alert-info">
  **Wait!** How do you do a line plot? I'm not telling you, but there a are two hints. Try running `?plot`   on the R console and try searching on Google!
</div>


```{r plot, exercise=TRUE}
x = seq(0, 10, 0.01)
y = 2 * sin(2 * x) + cos(x) + 3 * cos(3 * x)

```

```{r plot-solution}
x = seq(0, 10, 0.01)
y = 2 * sin(2 * x) + cos(x) + 3 * cos(3 * x)
plot(x, y, 'l')
```

### Installing pakages

The plot was good, but `ggplot2` is better. Let us install `ggplot2` first. The function `install.packages` installs packages from the CRAN archive.

```{r ggplot-install, exercise=TRUE}



```

```{r ggplot-install-solution}
install.packages('ggplot2')
```

### Better plotting

Try and create the same plot as before using `ggplot2`. Remember to load the package with `library` before using it.

```{r ggplot, exercise=TRUE}
# load package with library

# create some data
x = seq(0, 10, 0.01)
y = 2 * sin(2 * x) + cos(x) + 3 * cos(3 * x)

# make plot

```

```{r ggplot-solution}
# load package with library
library(ggplot2)

# create some data
x = seq(0, 10, 0.01)
y = 2 * sin(2 * x) + cos(x) + 3 * cos(3 * x)

# make plot
qplot(x, y, geom = 'line')
```


