---
title: "Meeting R"
output:
  learnr::tutorial:
      progressive: true
runtime: shiny_prerendered
description: >
  Learn the basics of R: interpreter, expressions, evaluation and variables.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Functions

Functions are reusable chunks of code which perform a specific operation. Many useful functions are defined in existing R packages, but you can always write your own.


### Base functions

Many useful functions have already been implemented. No need to reinvent the wheel! Try printing the variables in the following chunk using the `print` function of the `base` package.

```{r print, exercise=TRUE}
a = 1
b = c(1.1, 2.2, 3.3)
l = list(a='miao', b=2)

```

```{r print-solution}
a = 1
b = c(1.1, 2.2, 3.3)
l = list(a='miao', b=2)

print(a)
print(b)
print(l)
```

### Plotting

Also, R has great functions for plotting. Use the `plot` function in the `base` package to perform a line plot of x versus y.

<div class="alert alert-info">
  **Wait!** How do you do a line plot? I'm not telling you, but there a are two hints. Try running `?plot`   on the R console and try searching on Google!
</div>


```{r plot, exercise=TRUE}
x = seq(0, 10, 0.01)
y = 2 * sin(2 * x) + cos(x) + 3 * cos(3 * x)

```

```{r plot-solution}
x = seq(0, 10, 0.01)
y = 2 * sin(2 * x) + cos(x) + 3 * cos(3 * x)
plot(x, y, 'l')
```

### Installing pakages

The plot was good, but `ggplot2` is better. Let us install `ggplot2` first. The function `install.packages` installs packages from the CRAN archive.

```{r ggplot-install, exercise=TRUE}



```

```{r ggplot-install-solution}
install.packages('ggplot2')
```

### Better plotting

Try and create the same plot as before using `ggplot2`. Remember to load the package with `library` before using it.

```{r ggplot, exercise=TRUE}
# load package with library

# create some data
x = seq(0, 10, 0.01)
y = 2 * sin(2 * x) + cos(x) + 3 * cos(3 * x)

# make plot

```

```{r ggplot-solution}
# load package with library
library(ggplot2)

# create some data
x = seq(0, 10, 0.01)
y = 2 * sin(2 * x) + cos(x) + 3 * cos(3 * x)

# make plot
qplot(x, y, geom = 'line')
```

## Writing functions

Writing functions is a daily task for a developer. We'll start with some easy example here. We'll go into more details in the next sessions.

### Summing numbers

Write a simple function to sum two numbers and test it.

```{r fun1, exercise=TRUE}
add_numbers = function(a, b) {
  # complete here
}

add_numbers(2, 2)
```

```{r fun1-solution}
add_numbers = function(a, b) {
  return(a + b)
}

add_numbers(2, 2)
```

### Testing corner cases

What happens if we try to sum two strings? Let us try. Use the function `add_numbers` on two variables of type `character`.

```{r fun2, exercise=TRUE}





```

```{r fun2-solution}
add_numbers = function(a, b) {
  return(a + b)
}

x = 'miao'
y = 'bau'

add_numbers(x, y)
```

### Adding type checks

It does not work. In other languages it would, but R does not allow you to use the `+` operator on strings. So, maybe we should add some type of check. Modify the function `add_numbers` so that it throws an error if the type of either input is not numeric. Then, test it with strings.

<div class="alert alert-info">
  **Wait!** What do you mean with "throwing errors"? Try running `?stop` on the R console and Google for the `if` statement in R!
</div>

```{r fun3, exercise=TRUE}
add_numbers = function(a, b) {
  # complete here
}

x = 'miao'
y = 'bau'

add_numbers(x, y)
```

```{r fun3-solution}
add_numbers = function(a, b) {
  if ((!is.numeric(a)) | (!is.numeric(b))){
    stop('You shall not pass. Only numeric inputs are allowed.')
  }
  return(a + b)
}

x = 'miao'
y = 'bau'

add_numbers(x, y)
```

### Managing strings

Last update. We need now to make a new function `add_stuff`, with the same behavior of `add_numbers` for numeric inputs but capable of handling strings. If two strings are passed, we must return the concatenation of the strings. For example, `add_stuff('bau', 'miao')` should return `[1] 'baumiao'`.

<div class="alert alert-info">
  **Hint** Have a look to `?paste0` and Google for the `if` statement in R!
</div>

```{r fun4, exercise=TRUE}
add_stuff = function(a, b) {
  # complete here
}

print(add_stuff('bau', 'miao'))
print(add_stuff(2, 2))
print(add_stuff(2, 'miao')) # this line should result in an error. Remove it in the exercise to see the difference
```

```{r fun4-solution}
add_stuff = function(a, b) {
  if (is.numeric(a) & is.numeric(b)){
    return(a + b)
  } else if (is.character(a) & is.character(b)) {
    return(paste0(a, b))
  }
  stop('You shall not pass. Only numeric inputs are allowed.')
}

print(add_stuff('bau', 'miao'))
print(add_stuff(2, 2))
print(add_stuff(2, 'miao')) # this line should result in an error. Remove it in the exercise to see the difference
```
